### 2020-05-23 ms/main.tex
added used packages to the DESCRIPTION

\documentclass{article}

%% Language and font encodings
\usepackage[english]{babel}
\usepackage[utf8x]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{verbatim} % multiline comments

%% Sets page size and margins
\usepackage[a4paper,top=3cm,bottom=2cm,left=3cm,right=3cm,marginparwidth=1.75cm]{geometry}

%% Useful packages
\usepackage{amsmath}
\usepackage{float}
\usepackage{rotating}
\usepackage{booktabs}
%\restylefloat{figure} % for force placing figures
%\floatplacement{figure}{H}
%\floatplacement{table}{H}
%\floatplacement{longtable}{H}
%\usepackage[markers,nolists]{endfloat}
\usepackage{graphicx}
\usepackage[colorinlistoftodos]{todonotes}
\usepackage[colorlinks=true, allcolors=blue]{hyperref}
%\usepackage{apalike}
\usepackage[round]{natbib}
\usepackage[switch]{lineno} % line numbers
\usepackage{longtable}
\usepackage{multirow}
% Supp mat
\newcommand{\beginsupplement}{%
        \setcounter{table}{0}
        \renewcommand{\thetable}{S\arabic{table}}%
        \setcounter{figure}{0}
        \renewcommand{\thefigure}{S\arabic{figure}}%
     }

\title{Dewlap color variation in \textit{Anolis sagrei} is maintained between habitats within islands of the West Indies}
\author{
    \textsc{Rapha\"{e}l Scherrer$^{1,3}$ \thanks{Corresponding author: r.scherrer@rug.nl}, Colin M. Donihue$^{1,4}$, }\\
	\textsc{R. Graham Reynolds$^2$, Jonathan B. Losos$^{1,4}$ and Anthony J. Geneva$^1$} \\[1ex]
	%\thanks{A thank you} \\[1ex] % Your name
	\normalsize $^1$ Department of Organismic and Evolutionary Biology and Museum of Comparative Zoology \\ \normalsize Harvard University, Cambridge, MA, USA \\
	\normalsize $^2$ Department of Biology, University of North Carolina Asheville, Asheville, NC, USA\\ % Your institutionhttps://www.overleaf.com/project/5a0705e5f4573c1ace38dbdf
	\normalsize $^3$ Current address: Groningen Institute for Evolutionary Life Sciences,\\
	\normalsize Groningen, The Netherlands\\
	\normalsize $^4$ Current address: Department of Biology, Washington University, St. Louis, MO, USA\\
}
\date{} % Leave empty to omit a date

\begin{document}

\linenumbers

\maketitle

\begin{abstract}
    \input{sections/abstract.tex}
\end{abstract}

\textbf{Keywords} --- \textit{Anolis}, reflectance, local adaptation, sexual signal, supervised machine learning

\section*{Introduction}

\begin{table}
	\input{../analyses/01-counts/table_counts.tex}
\end{table}

\input{sections/introduction.tex}

\section*{Methods}

\input{sections/methods.tex}

\section*{Results}

\input{sections/results.tex}

\section*{Discussion}

\input{sections/discussion.tex}

\subsection*{Acknowledgements}

\input{sections/acknowledgments.tex}

\pagebreak

\section*{Figures}

\input{sections/figures.tex}

\pagebreak

\section*{Tables}

\input{sections/tables.tex}

\pagebreak

\bibliographystyle{apalike}
\bibliography{library.bib}

\beginsupplement

%\section*{Supplementary Figures}

%\input{sections/suppfigures.tex}

\pagebreak

\section*{Supplementary Figures}

\input{sections/suppfigures}

\pagebreak

\section*{Supplementary Tables}

\input{sections/supptables.tex}

\end{document}