### 2020-05-23 ms/sections/acknowledgments.tex

Collection permission was granted by the Bahamas Environment, Science and Technology Commission, the Bahamas National Trust, and the Cayman Islands Department of the Environment. The authors thank Sofia Prado-Irwin, Pavitra Muralidhar, Nicholas Herrmann, Richard E. Glor, Alberto R. Puente-Rol\'{o}n, Kevin Aviles-Rodriguez, Kristin Winchell, and Melissa Kemp for assistance in the field. Funding for this work was provided by the Templeton Foundation (to JBL), NSF DEB~\#1500761 (to AJG), NSF DBI~\#1609284 (to CMD), and a Harvard Museum of Comparative Zoology Putnam Grant (to RGR).