### 2020-05-23 ms/tables/table_anova.tex

% ANOVA table
\begin{table}[H]
    \centering
    \caption{Analyses of variance (ANOVA) testing the effect of habitat-type on multiple response variables across the most significant islands according to the SVM classification (Figure \ref{fig:classif_svm_pca}). Best fit refers to the fitting method, i.e. OLS, ordinary least squares (favored when $\Delta$AICc $> 0$) or GLS, generalized least squares (favored when $\Delta$AICc $< 0$). AICc values are presented for the best-fitting model. df, degrees of freedom of the model; df$_{\mbox{LRT}}$, degrees of freedom of the likelihood ratio test (LRT) used to test the habitat effect; Log-lik., log-likelihood of the full model, i.e. including the habitat effect; $\chi^2$, likelihood-ratio between the full model and a null model excluding the habitat effect. *, P < 0.05; **, P < 0.01; ***, P < 0.001.}
    \begin{tabular}{lllrrrrrrll}
        \hline
        Island & Variable & Best fit & df & AICc & $\Delta$AICc & df$_{\mbox{LRT}}$ & Log-lik. & $\chi^2$ & P-value & \\
        \hline
        Abaco & PC1 & OLS & 4 & 710.40 & 2.16 & 2 & -356.96 & 0.14 & 0.9308 & \\
        Abaco & PC2 & OLS & 4 & 620.06 & 4.02 & 2 & -310.16 & 31.74 & < 0.0001 &  ***\\
        Abaco & PC3 & OLS & 4 & 517.80 & 2.01 & 2 & -257.18 & 27.37 & < 0.0001 &  ***\\
        Abaco & PC4 & OLS & 4 & 440.59 & 0.78 & 2 & -217.18 & 1.36 & 0.5070 &  \\
        Bimini & PC1 & OLS & 4 & 561.34 & 0.77 & 2 & -283.06 & 7.40 & 0.0248 &  *\\
        Bimini & PC2 & OLS & 4 & 448.06 & 1.29 & 2 & -223.77 & 8.09 & 0.0175 &  *\\
        Bimini & PC3 & GLS & 6 & 405.30 & -0.23 & 2 & -199.18 & 10.39 & 0.0056 &  **\\
        Bimini & PC4 & OLS & 4 & 274.15 & 3.54 & 2 & -132.74 & 0.33 & 0.8499 &  \\
        Cayman Brac & PC1 & GLS & 6 & 402.81 & -4.05 & 2 & -200.86 & 13.81 & 0.0010 &  **\\
        Cayman Brac & PC2 & OLS & 4 & 332.14 & 3.51 & 2 & -165.91 & 8.41 & 0.0149 &  *\\
        Cayman Brac & PC3 & OLS & 4 & 295.78 & 2.77 & 2 & -146.57 & 27.16 & < 0.0001 &  ***\\
        Cayman Brac & PC4 & OLS & 4 & 279.24 & 4.33 & 2 & -137.77 & 5.63 & 0.0600 &  \\
        Little Cayman & PC1 & OLS & 4 & 367.17 & 2.50 & 2 & -185.99 & 8.18 & 0.0167 &  *\\
        Little Cayman & PC2 & GLS & 6 & 287.61 & -3.61 & 2 & -140.46 & 29.76 & < 0.0001 &  ***\\
        Little Cayman & PC3 & OLS & 4 & 277.69 & 1.41 & 2 & -138.06 & 21.34 & < 0.0001 &  ***\\
        Little Cayman & PC4 & OLS & 4 & 226.70 & 2.53 & 2 & -110.74 & 2.85 & 0.2410 &  \\
        Long Island & PC1 & GLS & 6 & 442.34 & -2.09 & 2 & -221.16 & 2.91 & 0.2331 &  \\
        Long Island & PC2 & GLS & 6 & 351.37 & -3.08 & 2 & -172.57 & 4.52 & 0.1043 &  \\
        Long Island & PC3 & OLS & 4 & 322.06 & 3.67 & 2 & -159.98 & 11.24 & 0.0036 &  **\\
        Long Island & PC4 & OLS & 4 & 195.45 & 2.38 & 2 & -92.88 & 6.46 & 0.0395 &  *\\
        \hline
    \end{tabular}
    \label{tab:anova}
\end{table}