### 2020-05-24 ms/sections/tables.tex

\begin{table}
	\label{tab:anova}
	\caption{}
	\centering
	\input{"../analyses/07-ANOVA/table_anova.tex"}
\end{table}