### 2020-05-24 analyses/07-ANOVA/table_posthoc.tex
regenerated tables

\begin{tabular}{lllllrrl}
\toprule
Island & Variable & Test & Contrast &  & Statistic & $P$ & \\
\midrule
Abaco & PC2 & tukey & coastal & coppice & -0.71 & 0.8690 & \\
Abaco & PC2 & tukey & coastal & mangrove & -8.28 & 0.0000 & ***\\
Abaco & PC2 & tukey & coppice & mangrove & -6.82 & 0.0000 & ***\\
Abaco & PC3 & tukey & coastal & coppice & 4.89 & 0.0025 & **\\
Abaco & PC3 & tukey & coastal & mangrove & 7.52 & 0.0000 & ***\\
Abaco & PC3 & tukey & coppice & mangrove & 2.55 & 0.1746 & \\
Bimini & PC1 & tukey & coastal & coppice & -2.57 & 0.1726 & \\
Bimini & PC1 & tukey & coastal & mangrove & -3.49 & 0.0428 & *\\
Bimini & PC1 & tukey & coppice & mangrove & -0.70 & 0.8741 & \\
Bimini & PC2 & tukey & coastal & coppice & 2.10 & 0.3044 & \\
Bimini & PC2 & tukey & coastal & mangrove & 3.91 & 0.0200 & *\\
Bimini & PC2 & tukey & coppice & mangrove & 1.44 & 0.5674 & \\
Bimini & PC3 & dunnett & coastal & coppice & 3.59 & 0.0065 & **\\
Bimini & PC3 & dunnett & coastal & mangrove & -0.21 & 0.9954 & \\
Bimini & PC3 & dunnett & coppice & mangrove & -3.03 & 0.0162 & *\\
Cayman Brac & PC1 & dunnett & coastal & coppice & -2.93 & 0.0225 & *\\
Cayman Brac & PC1 & dunnett & coastal & mangrove & -0.74 & 0.8436 & \\
Cayman Brac & PC1 & dunnett & coppice & mangrove & 3.60 & 0.0032 & **\\
Cayman Brac & PC2 & tukey & coastal & coppice & 2.28 & 0.2491 & \\
Cayman Brac & PC2 & tukey & coastal & mangrove & 4.15 & 0.0140 & *\\
Cayman Brac & PC2 & tukey & coppice & mangrove & 1.98 & 0.3476 & \\
Cayman Brac & PC3 & tukey & coastal & coppice & 3.46 & 0.0467 & *\\
Cayman Brac & PC3 & tukey & coastal & mangrove & 8.16 & 0.0000 & ***\\
Cayman Brac & PC3 & tukey & coppice & mangrove & 4.97 & 0.0028 & **\\
Little Cayman & PC1 & tukey & coastal & coppice & -4.08 & 0.0167 & *\\
Little Cayman & PC1 & tukey & coastal & mangrove & -1.51 & 0.5389 & \\
Little Cayman & PC1 & tukey & coppice & mangrove & 2.65 & 0.1593 & \\
Little Cayman & PC2 & dunnett & coastal & coppice & 0.88 & 0.7637 & \\
Little Cayman & PC2 & dunnett & coastal & mangrove & -4.02 & 0.0014 & **\\
Little Cayman & PC2 & dunnett & coppice & mangrove & -6.99 & 0.0000 & ***\\
Little Cayman & PC3 & tukey & coastal & coppice & -4.28 & 0.0114 & *\\
Little Cayman & PC3 & tukey & coastal & mangrove & -7.03 & 0.0000 & ***\\
Little Cayman & PC3 & tukey & coppice & mangrove & -2.18 & 0.2822 & \\
Long Island & PC3 & tukey & coastal & coppice & 3.75 & 0.0285 & *\\
Long Island & PC3 & tukey & coastal & mangrove & 4.19 & 0.0127 & *\\
Long Island & PC3 & tukey & coppice & mangrove & 0.47 & 0.9408 & \\
Long Island & PC4 & tukey & coastal & coppice & 0.29 & 0.9766 & \\
Long Island & PC4 & tukey & coastal & mangrove & 3.47 & 0.0454 & *\\
Long Island & PC4 & tukey & coppice & mangrove & 2.81 & 0.1257 & \\
\bottomrule
\end{tabular}