### 2020-05-24 analyses/07-ANOVA/table_postkw.tex

\begin{tabular}{lllllrrl}
\toprule
Island & Variable & Test & Contrast &  & Statistic & $P$ & \\
\midrule
Abaco & PC2 & nemenyi & coastal & coppice & 0.43 & 0.9511 & \\
Abaco & PC2 & nemenyi & coastal & mangrove & 6.50 & 0.0000 & ***\\
Abaco & PC2 & nemenyi & coppice & mangrove & 5.47 & 0.0003 & ***\\
Bimini & PC1 & nemenyi & coastal & coppice & 2.62 & 0.1523 & \\
Bimini & PC1 & nemenyi & coastal & mangrove & 3.42 & 0.0411 & *\\
Bimini & PC1 & nemenyi & coppice & mangrove & 0.60 & 0.9048 & \\
Bimini & PC3 & nemenyi & coastal & coppice & 4.97 & 0.0013 & **\\
Bimini & PC3 & nemenyi & coastal & mangrove & 0.92 & 0.7924 & \\
Bimini & PC3 & nemenyi & coppice & mangrove & 4.93 & 0.0014 & **\\
Little Cayman & PC3 & nemenyi & coastal & coppice & 3.73 & 0.0227 & *\\
Little Cayman & PC3 & nemenyi & coastal & mangrove & 6.23 & 0.0000 & ***\\
Little Cayman & PC3 & nemenyi & coppice & mangrove & 2.00 & 0.3345 & \\
Long Island & PC1 & nemenyi & coastal & coppice & 4.31 & 0.0065 & **\\
Long Island & PC1 & nemenyi & coastal & mangrove & 0.26 & 0.9816 & \\
Long Island & PC1 & nemenyi & coppice & mangrove & 3.94 & 0.0148 & *\\
\bottomrule
\end{tabular}