### 2020-05-24 ms/sections/tables.tex

\begin{sidewaystable}
	\label{tab:anova}
	\caption{Significance of habitat differences in dewlap coloration, using ANOVA for all islands where significant multivariate differences in dewlap coloration were detected by SVMs. Best fitting model: 1, OLS; 2, GLS. df, degrees of freedom. $\Delta$AICc, difference in AICc between the best fitting model and the OLS-model. AICcw, AICc weight. LRT, likelihood ratio test. Log-lik., log-likelihood. $\chi^2$, likelihood ratio. *, P < 0.05; **, P < 0.01, ***, P < 0.001.}
	\centering
	\input{"../analyses/07-ANOVA/table_anova.tex"}
\end{sidewaystable}