### 2021-05-29 ms/sections/figures.tex
island maps and island figures refurbished

\begin{figure}[H]
	\centering
	\includegraphics[width=\textwidth]{figures/Abaco_2.2.png}
	\caption{Comparison of dewlap coloration across habitats on Abaco. (A) Map of the island with the sampling sites colored by habitat. (B) Reflectance profiles of all the dewlaps on the island. (C) How reflectance profiles map onto the within-island principal components. (D) Confusion matrix showing the proportion of lizards from each (true) habitat reassigned to each (predicted) habitat by the random forests, based on the first four within-island principal components and averaged across replicates. Each column sums to one. (E) Within-island principal component scores across habitats. Bars indicate significant contrasts. *, $P < 0.05$; **, $P < 0.01$; ***, $P < 0.001$.}
	\label{fig:Abaco}
\end{figure}