### 2021-09-06 ms/main_title.tex
stuff

\documentclass{article}

% General rule for P-values
% * < 0.05
% ** < 0.01
% *** < 0.001
% Write "< 0.0001" or the value up to the fourth decimal

%% Language and font encodings
\usepackage[english]{babel}
\usepackage[utf8x]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{verbatim} % multiline comments

%% Sets page size and margins
\usepackage[a4paper,top=3cm,bottom=2cm,left=3cm,right=3cm,marginparwidth=1.75cm]{geometry}

%% Useful packages
\usepackage{amsmath}
\usepackage{float}
\usepackage{rotating}
\usepackage{booktabs}
%\restylefloat{figure} % for force placing figures
%\floatplacement{figure}{H}
%\floatplacement{table}{H}
%\floatplacement{longtable}{H}
%\usepackage[markers,nolists]{endfloat}
\usepackage{graphicx}
\usepackage[colorinlistoftodos]{todonotes}
\usepackage[colorlinks=true, allcolors=blue]{hyperref}
%\usepackage{apalike}
\usepackage[round]{natbib}
\usepackage[switch]{lineno} % line numbers
\usepackage{longtable}
\usepackage{multirow}
% Supp mat
\newcommand{\beginsupplement}{%
	\setcounter{table}{0}
	\renewcommand{\thetable}{S\arabic{table}}%
	\setcounter{figure}{0}
	\renewcommand{\thefigure}{S\arabic{figure}}%
}

\title{Dewlap color variation in \textit{Anolis sagrei} is maintained among habitats within islands of the West Indies}
\author{
	\textsc{Rapha\"{e}l Scherrer}$^{1,3}$ \thanks{Corresponding author} (raphael.scherrer@evobio.eu, ORCID 0000-0002-1447-7630),\\
    \textsc{Colin M. Donihue}$^{1,4}$ (colindonihue@gmail.com, 0000-0003-1096-8536),\\
	\textsc{R. Graham Reynolds}$^2$ (greynold@unca.edu, 0000-0002-0056-8304),\\ 
    \textsc{Jonathan B. Losos}$^{1,4}$ (losos@wustl.edu, 0000-0003-4712-8258) and\\
	\textsc{Anthony J. Geneva}$^{1,5}$ (anthony.geneva@gmail.com, 0000-0001-8253-6527)\\[1ex]
	\normalsize $^1$ Department of Organismic and Evolutionary Biology and Museum of Comparative Zoology \\ \normalsize Harvard University, Cambridge, MA, USA \\
	\normalsize $^2$ Department of Biology, University of North Carolina Asheville, Asheville, NC, USA\\ 
	\normalsize $^3$ Current address: Groningen Institute for Evolutionary Life Sciences,\\
	\normalsize University of Groningen, Groningen, The Netherlands\\
	\normalsize $^4$ Current address: Department of Biology, Washington University, St. Louis, MO, USA\\
	\normalsize $^5$ Current address: Department of Biology, Center for Computational and Integrative Biology,\\ \normalsize Rutgers University--Camden, Camden, NJ, USA
}
\date{} % Leave empty to omit a date

\begin{document}

	\linenumbers

	\maketitle

	\textbf{Running title} --- Dewlap coloration in \textit{Anolis sagrei}.

\end{document}