### 2021-09-06 ms/main_title.tex

\documentclass{article}

%% Language and font encodings
\usepackage[english]{babel}
\usepackage[utf8x]{inputenc}
\usepackage[T1]{fontenc}

%% Sets page size and margins
\usepackage[a4paper,top=3cm,bottom=2cm,left=3cm,right=3cm,marginparwidth=1.75cm]{geometry}

\title{Dewlap color variation in \textit{Anolis sagrei} is maintained among habitats within islands of the West Indies}

\author{

	\textsc{Rapha\"{e}l Scherrer}$^{1,3}$ \thanks{Corresponding author} (raphael.scherrer@evobio.eu, ORCID 0000-0002-1447-7630),\\
    \textsc{Colin M. Donihue}$^{1,4}$ (colindonihue@gmail.com, 0000-0003-1096-8536),\\
	\textsc{R. Graham Reynolds}$^2$ (greynold@unca.edu, 0000-0002-0056-8304),\\ 
    \textsc{Jonathan B. Losos}$^{1,4}$ (losos@wustl.edu, 0000-0003-4712-8258) and\\
	\textsc{Anthony J. Geneva}$^{1,5}$ (anthony.geneva@gmail.com, 0000-0001-8253-6527)\\[1ex]

	\normalsize $^1$ Department of Organismic and Evolutionary Biology and Museum of Comparative Zoology \\ \normalsize Harvard University, Cambridge, MA, USA \\
	\normalsize $^2$ Department of Biology, University of North Carolina Asheville, Asheville, NC, USA\\ 
	\normalsize $^3$ Current address: Groningen Institute for Evolutionary Life Sciences,\\
	\normalsize University of Groningen, Groningen, The Netherlands\\
	\normalsize $^4$ Current address: Department of Biology, Washington University, St. Louis, MO, USA\\
	\normalsize $^5$ Current address: Department of Biology, Center for Computational and Integrative Biology,\\ \normalsize Rutgers University--Camden, Camden, NJ, USA

}

\date{} % Leave empty to omit a date

\begin{document}

	\maketitle

	\textbf{Running title} --- Dewlap coloration in \textit{Anolis sagrei}.

\end{document}