### 2022-03-09 ms/final/final/main.tex

\documentclass{article}

% Language and font encodings
\usepackage[english]{babel}
\usepackage[utf8x]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{verbatim} % multiline comments

% Sets page size and margins
\usepackage[a4paper,top=3cm,bottom=2cm,left=3cm,right=3cm,marginparwidth=1.75cm]{geometry}

% Useful packages
\usepackage{amsmath}
\usepackage{float}
\usepackage{rotating}
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage[colorinlistoftodos]{todonotes}
\usepackage[colorlinks=true, allcolors=blue]{hyperref}
\usepackage[round]{natbib}
\usepackage[switch]{lineno} % line numbers
\usepackage{longtable}
\usepackage{multirow}

% Supplementary Materials
\newcommand{\beginsupplement}{
	\setcounter{table}{0}
    \renewcommand{\thetable}{S\arabic{table}}
    \setcounter{figure}{0}
    \renewcommand{\thefigure}{S\arabic{figure}}
}

\title{Dewlap color variation in \textit{Anolis sagrei} is maintained among habitats within islands of the West Indies}

\author{

    \textsc{Rapha\"{e}l Scherrer$^{1,3}$ \thanks{Corresponding author: r.scherrer@rug.nl}, Colin M. Donihue$^{1,4}$, }\\
	\textsc{R. Graham Reynolds$^2$, Jonathan B. Losos$^{1,4}$ and Anthony J. Geneva$^{1,5}$} \\[1ex]

	\normalsize $^1$ Department of Organismic and Evolutionary Biology and Museum of Comparative Zoology \\ \normalsize Harvard University, Cambridge, MA, USA \\
	\normalsize $^2$ Department of Biology, University of North Carolina Asheville, Asheville, NC, USA\\ 
	\normalsize $^3$ Current address: Groningen Institute for Evolutionary Life Sciences,\\
	\normalsize University of Groningen, Groningen, The Netherlands\\
	\normalsize $^4$ Current address: Department of Biology, Washington University, St. Louis, MO, USA\\
	\normalsize $^5$ Current address: Department of Biology, Center for Computational and Integrative Biology,\\ \normalsize Rutgers University--Camden, Camden, NJ, USA

}

\date{} % Leave empty to omit a date

\begin{document}

	\linenumbers

	\maketitle

	\begin{abstract}
    	\input{sections/abstract.tex}
	\end{abstract}

	\textbf{Keywords} --- reflectance, adaptation, sexual signal, machine learning, polymorphism

	\section*{Introduction}

	\input{sections/introduction.tex}

	\pagebreak

	\section*{Methods}

	\input{sections/methods.tex}

	\pagebreak

	\section*{Results}

	\input{sections/results.tex}

	\pagebreak

	\section*{Discussion}

	\input{sections/discussion.tex}

	\pagebreak

	\section*{Acknowledgements}

	\input{sections/acknowledgments.tex}

	\pagebreak

	\section*{Figures}

	\input{sections/figures.tex}

	\pagebreak

	\section*{Tables}

	\input{sections/tables.tex}

	\pagebreak

	\bibliographystyle{apalike}
	\bibliography{library.bib}

	\beginsupplement

	\pagebreak

	\section*{Appendix}

	\input{sections/appendix}

	\pagebreak

	\section*{Supplementary Figures}

	\input{sections/suppfigures}

	\pagebreak

	\section*{Supplementary Tables}

	\input{sections/supptables.tex}

\end{document}